\documentclass{master_thesis}

\author{Mykyta Sikriier}
\studentbooknumber{381815}

\title{
    Commercial Banking Front Office: new approach to serve customers
}
\titlepl{
    Front Office Bankowości Komercyjnej: nowe podejście do obsługi klientów
}

\course{Quantitative Finance}
\studycode{Economics (14300)}

\supervisor{PhD Krzysztof Spirzewski}
\supervisorunit{Finance and Accounting}

\city{Warsaw}
\date{September 2021}

\classification{}

\keywords{
    credit institutions, information technologies, virtual banking, chat-bots
}

\summary{
    In 2007 PSD had shown an interest of financial regulators in banking competitiveness by digital instruments.
    Later, PSD2 triggered a new wave of banking digitalization, openness, Big Data usage and third-party integrations.
    This thesis seeks to analyze the ways of development of commercial banks Front Offices using ML \& AI technologies.
    The research focuses on the importance of openness brought by PSD2 and Open Banking initiative.
    Study investigates ways of application of ML \& AI technologies on various layers of a bank and targets the importance of usage of available customer data in order to achieve market efficiency.
    The thesis proposes possible Chat-Bot solutions as a Front-office, as a Middle-office kit for Front-office employees, as a product, which uses Banking-as-a-Platform and as a customer service solution for existing Banking-as-a-Service.
}

\summarypl{
    W 2007 roku PSD pokazał zainteresowanie regulatorów rynku finansowego w konkurencyjności bankowej za pomocą instrumentów cyfrowych.
    Później, PSD2 spowodował falę cyfryzacji bankowej, otwartości, wykorzystania Big Data i integracji z firmami zewnętrznymi.
    Niniejsza praca ma na celu przeanalizować drogi rozwoju Front Officów banków komercyjnych za pomocą technologii ML \& AI.
    Badania koncentrują się wokół znaczenia otwartości, wniesioną przez PSD2 i inicjatywy Open Banking.
    Praca bada sposoby na wykorzystania z technologii ML \& AI na różnych poziomach banku i ocenia znaczenia wykorzystania istniejących danych użytkowników w celu osiągnięcia efektywności rynkowej.
    Niniejsza praca proponuje możliwe rozwiązania jako rozwiązanie Chat-botowe dla Front-office, jako Middle-office zestaw dla pracowników Front-office, jako produkt w ramach Banking-as-a-Platform i jako rozwiązanie do obsługi klientów dla istniejącego Banking-as-a-Service.
}

\begin{document}

\maketitle

\makesummary

% List
\tableofcontents

% Introduction
\input{content/introduction.tex}

% Chapter 1
\input{content/chapter_1/chapter_1.tex}

% Chapter 2
\input{content/chapter_2/chapter_2.tex}

% Chapter 3
\input{content/chapter_3/chapter_3.tex}

% Conclusion
\input{content/conclusions.tex}

% Literature
\input{content/bibliography.tex}

% Appendices
\input{content/appendices.tex}

% Annexes
\input{content/annexes.tex}

\end{document}
