% !TeX root = ../../master_thesis.tex

\chapter{Artificial Intelligence in Commercial Banking} \label{ch:ml_ai}

%\section*{Introduction}
Most of the work in banking industry, which is not connected to human interaction, is over-documented and formalized. 
Having sufficient amount of input data and knowing what kind of output is required it is possible to create a formalized determined way of how to obtain output from input. 
This determined formalized way is called an algorithm.

Main problem of common algorithms is, as it comes of definition, is that it is impossible for an algorithm to solve problems, for which it is not designed.

Nevertheless, it is possible to create an algorithm, that can adapt to changing environment based on previous results.
To find out how to use collect, process and analyze existing data of previous states of environment and corresponding result we can use various statistical instruments.
Aside general statistics, of course, one is able to use tools of data science and machine learning.

As a result, it is possible to create a system, which would be able taking into account primary input data and its own past output data produce new adapted output data corresponding to ever-changing environment.
Thus, having a system which is able to serve decisions, which were not available or known before, we can call this system intelligent.
As a result, we could tell that we had an Artificial Intelligence — AI.
According to European Parliament, Artificial Intelligence refer to systems that display intelligent behavior by analyzing their environment and taking action with some degree of autonomy in order to achieve specific goals.
\cite{ai_ep_definition}
AI is typically defined as the ability of a machine to perform cognitive functions we associate with human minds, such as perceiving, reasoning, learning, and problem-solving. 
Examples of technologies that enable AI to solve business problems are robotics and autonomous vehicles, computer vision, language, virtual agents, and machine learning.
\cite{executive_guide_to_ai}


\input{content/chapter_2/1_ai_history.tex}
\input{content/chapter_2/2_ai_general.tex}


%\section*{Conclusions}
Artificial Intelligence has already become a leading driver for technology innovations in banking sector.
It has already been developed enough and definitely reliable in everything regarding risks, confidentiality, human factor and marketing strategies.

Nevertheless, AI has already changed financial business and will change it even more in near future.
The vast majority of actions, that are being done both in back-office, middle-office and front-office by employees can be automized and algorithmized.

\begin{table}
    \centering
    \begin{tabular}{| c | c | c | c |}
        \hline 
        \textbf{Divisions} &
            Front office & 
            Middle office & 
            Back office \\ \hline 
       
        \textbf{More mature} & 
            Conversational banking & 
            Anti-fraud, KYC, AML & 
            Credit underwriting \\ \hline 
       
        \textbf{Less mature} & 
            Biometrics & 
            Legal and Compliance &
            Smart contracts \\ \hline 
        \textbf{Savings} &
            \$199 billion &
            \$217 billion &
            \$32 billion \\
        \hline 
    \end{tabular}
    \caption{Fintech AI Use-cases}
    \medskip
    \footnotesize\textit{Source:} Own study, based on: Lex Sokolin, Matt Low, "\#Machine Intelligence \& Augmented Finance", Autonomous Research, p. 29.
\end{table}

The main interest for banks is in application of AI in front-office and middle-office, as it will lead to significant cost reduction, up to \$400 billion in 2025.
Banks use AI in front-office for identification and authentication of clients, in order to imitiate alive workers with chat-bots and voice assistants, to deepen relationships with customer, increase engagement, and to offer personalized ideas and recommendations.
Additionally, AI is being used in middle-office.
It allows detecting and preventing frauds with payments, upgrading processes connected to anti-money laundering processes and detection for know your customer control systems.
Those sources of income, which have already been transformed by AI, show the valid way to use this possibility.
Nevertheless, those strategies show that banks have to create holistic strategy, which would connect and cover multiple scopes of banking, sources of useful data, third-parties and employees.
\cite{autonomous_next}

However, with all the advantages of modern software, forecasts of full robotization probably won't be realized, as in near future it is unlikely for a program to replace a person.
\cite{deluxe_mid_size_banks_risk}

In order to integrate the foregoing into existing banking service it is necessary for banking platform to support it.
Obviously, those platforms, which current banking stands on were developed around 15-30 years ago, or even more, and this makes integration of modern algorithms almost impossible. 
This is why modern banks are now puzzled if it is possible to reengineer their software in order to have an effective possibility to embed Artificial Intelligence and Machine Learning in its banking platform.

Nevertheless, it is extremely important to remember, that automating workplaces doesn't always lead to dismissal of workers.
It is important to create retraining programs, that would allow existing employees to obtain new skills in fields,
that are out of risk of automation, in order to decrease automation's influence on labour market.

Increasing number of internet users, availability of smartphones and other handheld devices, and rapid development of mobile internet forms new users' habits and behaviors.
Social network users and mobile app users are more and more focused on getting instant results and target specific actions with the lowest amount of clicks possible.

However, the need of receiving high-quality financial services and personal consultations hasn't disappeared: people are still calling to call-centers.
Even though remote banking service systems are rapidly developing, number of phone calls recently significantly increased. 

Nevertheless, there are good news for banks that have yet to embark on their AI journey. 
Making progress in AI does not require an unbearable effort.
It typically begins with identifying your organizational goals, readying your data and finding the right partners.
Nonetheless, even medium-sized and small banks have to investigate this topic by now, as in future, gap between AI users and nonuser will only widen.





